---
title: "expdean tidy functions for batch modeling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{expdean tidy functions for batch modeling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(expdean)
```

## `lm_list()` function

```{r}
mt_trial <- mtcars %>%
  rownames_to_column("vehicle_name") %>%
  as_tibble() %>%
  mutate_at(c("vehicle_name", "vs", "am"), as.factor)

lm_model_list1 <- lm_list(mt_trial, c("mpg", "drat", "qsec"), c("am", "vs", "am:vs", "hp"))

```

## `model_mean_differences()` function

Linear mixed model can be fitted using `lme` package. Marginal and conditional means of continuous response variables can be extracted from this model class or `lmer` objects using estimated marginal means available in `emmeans` package.

```{r}
# pairwise comparison between treatment levels
pairs(emmeans::emmeans(lm_model_list1, "am"))
# p-value comparison
emmeans::pwpp(emmeans::emmeans(lm_model_list1, "am"))
```

Model contrasts and contrast coefficients can be assessed.

```{r}
# contrasts
emmeans::contrast(emmeans::emmeans(lm_model_list1, "am"))
coef(emmeans_contrast) # contrast coefficients can be assessed
```

Pairwise comparison of estimated marginal means:

```{r}
multcomp::cld(emmeans::emmeans(lm_model_list1, "am")) # compact display
```

