---
title: "Tidy genetics and tidy evaluation use"
author: "Deependra Dhakal"
date: "10/21/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

expr() and exprs() capture expressions that you supply:

```{r}
rlang::expr(symbol)
rlang::exprs(several, such, symbols)

```

enexpr() and enexprs() capture expressions that your user supplied:

```{r}
expr_inputs <- function(arg, ...) {
  user_exprs <- enexprs(arg, ...)
  user_exprs
}
expr_inputs(hello)
expr_inputs(hello, bonjour, ciao)
```

ensym() and ensyms() provide additional type checking to ensure the user calling your function has supplied bare object names:

```{r}
sym_inputs <- function(...) {
  user_symbols <- ensyms(...)
  user_symbols
}
sym_inputs(hello, "bonjour")
## sym_inputs(say(hello))  # Error: Must supply symbols or strings
expr_inputs(say(hello))
```

All these quoting functions have quasiquotation support. This means that you can unquote (evaluate and inline) part of the captured expression:

```{r}
what <- sym("bonjour")
expr(say(what))
expr(say(!!what))
```

This also applies to expressions supplied by the user. This is like an escape hatch that allows control over the captured expression:

```{r}
expr_inputs(say(!!what), !!what)
```

Finally, you can capture expressions as quosures. A quosure is an object that contains both the expression and its environment:

```{r}
quo <- quo(letters)
quo

rlang::get_expr(quo)
rlang::get_env(quo)
```

Quosures can be evaluated with eval_tidy():

```{r}
rlang::eval_tidy(quo)
```

They have the nice property that you can pass them around from context to context (that is, from function to function) and they still evaluate in their original environment:

```{r}
multiply_expr_by_10 <- function(expr) {
  # We capture the user expression and its environment:
  # expr <- enquo(expr)
  expr <- enexpr(expr)
  
  # Then create an object that only exists in this function:
  local_ten <- 10
  
  # Now let's create a multiplication expression that (a) inlines
  # the user expression as LHS (still wrapped in its quosure) and
  # (b) refers to the local object in the RHS:
  quo(!!expr * local_ten)
}
quo <- multiply_expr_by_10(2 + 3)
```

The local parts of the quosure are printed in colour if your terminal is capable of displaying colours:

```{r}
quo
```

All the quosures in the expression evaluate in their original context. The local objects are looked up properly and we get the expected result:

```{r}
rlang::eval_tidy(quo)
```


```{r}

simple_df <- readr::read_csv("./data/00 Sample RCBD data.csv", na = ".")
simple_df <- simple_df %>% 
  mutate_at(c("Country", "Location", "Rep", "Genotype"), as.factor)

heritability_n_blues <- function(df, grouping_var, grouping_items_sel,
                                 dependent_var, covariate = FALSE, genotype_var, replication_var){
  
  grvariable <- substitute(grouping_var) # capture the user expression and its environment
  gritems <- grouping_items_sel
  response <- substitute(dependent_var)
  co_variate <- substitute(covariate)
  genotype <- substitute(genotype_var) # capture expressions that your user supplied
  replication <- substitute(replication_var)
  response_exp <- paste0(quo_name(response), "_exp")
  co_variate_scaled <- paste0(quo_name(co_variate), "_scaled")
  
  # create working dataframe with response as expectation of the interested variable
  df_tmp <- df %>%
    dplyr::mutate(!! response_exp := (!!response-mean(!!response, na.rm = T))/stats::sd(!!response, na.rm = T)) %>%
    dplyr::mutate(!! co_variate_scaled := scale(!!co_variate, scale = T, center = T))
  
  # instantiate objects and parameters
  genotypic_var <- numeric()
  error_var <- numeric()
  h2 <- numeric()
  nRep <- df_tmp %>% dplyr::pull(!!replication) %>% unique() %>% length()

  models <- df_tmp %>%
    dplyr::filter(!!grvariable %in% grouping_items_sel) %>%
    dplyr::group_by(!!grvariable) %>%
    dplyr::summarise(model_rcbd_random = list(
      lme4::lmer(stats::reformulate(termlabels = c(paste("(1|", replication, ")"),
                                                   paste("(1|", genotype, ")"),
                                                   !! co_variate_scaled),
                                    response = !! response_exp))),
      model_rcbd_mixed = list(
        lme4::lmer(stats::reformulate(termlabels = c(paste("(1|", replication, ")"),
                                               genotype,
                                               !! co_variate_scaled), # to omit fixed effect covariate remove thismanually
                                response = !! response_exp)))
    )
    
  return(list(df_tmp, 
              models,
              co_variate_scaled, 
              response_exp, 
              replication, 
              genotype, 
              nRep))
  
}

hvalues_and_other <- heritability_n_blues(df = `simple_df`, grouping_var = `Location`, 
                     grouping_items_sel = c("Karaj", "Beijin", "Chihuahua"), 
                     dependent_var = `YLD`, genotype_var = `Genotype`, covariate = `TGW`,
                     replication_var = `Rep`)


hvalues_and_other
```

